Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago

### 2025-03-22 manuscript.Rmd
show suspected complexes as cf. spp.

![Representative photographs of the four most abundant crab species found across all deployed ARMS units (specimens shown were not collected as part of this study). (a) "*Chlorodiella laevissima*" (FLMNH 45490, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (b) "*Perinia tumida*" (FLMNH 45652, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (c) *Percnon abbreviatum* (FLMNH 45628, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (d) *Dynomene hispida* (FLMNH 69204, Manawai, Hawai&#x02BB;i). Photo credits: Florida Museum of Natural History, licensed [CC BY-NC 3.0](https://creativecommons.org/licenses/by-nc/3.0/)](data/images/4crabs.png){#fig:fig_crabs}

@@

ARMS units were retrieved by encapsulating them in-situ in a 100 μm mesh-lined crate to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate (see [@lerayDNABarcodingMetabarcoding2015] for processing details). All adult brachyuran crabs ($\geq$ 2 mm) were photographed, identified, preserved in 95% ethanol, and retained in a specimen archive at the Hawai&#x02BB;i Institute of Marine Biology. Specimens were initially identified to family or genus using Poore & Ahyong [-@pooreMarineDecapodCrustacea2023] and additionally to species where possible according to expert advice from Scott Godwin (former curator of invertebrate zoology, Bishop Museum) and Gustav Paulay (curator of invertebrate zoology, Florida Museum of Natural History). Where species-level identifications were not possible, taxa were collapsed to family or genus (depending on individual identification). We recognize that the species-level taxonomy of Indo-Pacific brachyurans remains poorly resolved, particularly for abundant groups like the Chlorodiellinae [@lasleyjrIntegrativeSystematicRevision2014; @lasleyjrPhylogeneticRelationshipsUbiquitous2015], and some "species" are known to comprise complexes [e.g. "*Perinia tumida*", @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009]. However, the prevalence of species complexes and cryptic diversity suggests that our assignments are likely to under- rather than overrepresent the true number of taxa. Thus, any significant patterns we detected likely persist in spite of the potential presence of cryptic diversity. Since our primary concern was categorical diversity of assemblages rather than accurate naming, we have retained species names in cases where complexes are suspected.

@@

```{r tbl-supp-species, results="asis", dependson="taxonomy-validation"}
# make table of all sampled taxa
species_tbl <- cc$crabs %>%
  psmelt() %>%
  # sum counts by island group, depth zone, and species
  group_by(OTU,island_group,hawaiian_name,depth_zone,depth,family,genus,species,complex) %>%
  summarise(n=sum(Abundance)) %>%
  ungroup() %>%
  # keep species we actually found
  filter(n > 0) %>%
  # figure out if a species is exclusive to a depth zone
  group_by(species) %>%
  mutate(zones = str_c(depth_zone,collapse="|")) %>%
  ungroup() %>%
  mutate(depth_exclusive = depth_zone == zones) %>%
  arrange(island_group,hawaiian_name,depth_zone,depth,family,genus,species) %>%
  # join in validated taxonomy
  left_join(crab_taxonomy_validated,by=c("OTU" = "otu")) %>%
  select(hawaiian_name,island_group,depth_zone,depth,family.y,display_species,depth_exclusive,complex=complex.y) %>%
  mutate(display_species = as.character(case_when(
    as.logical(complex) ~ str_glue('"{display_species}"'),
    .default = display_species
  ))) %>%
  select(-complex) %>%
  rename(island = hawaiian_name,valid_species=display_species) %>%
  mutate(
    across(c(island_group,depth_zone),str_to_sentence),
    depth_exclusive = if_else(depth_exclusive,"Yes","No")
  ) %>%
  rename_with(\(n) {
    n %>%
      str_replace_all("depth$","depth (m)") %>%
      str_replace_all("_"," ") %>%
      str_replace_all("\\.y$","") %>%
      str_replace_all("^display.","") %>%
      str_to_sentence()
  })
print_table(species_tbl,"tbl_supp_species","List of crab taxa detected by island and ARMS unit deployment depth",hdr_case="none",accuracy=1)
```

>>>>

![Representative photographs of the four most abundant crab species found across all deployed ARMS units (specimens shown were not collected as part of this study). (a) *Chlorodiella* cf. *laevissima* (FLMNH 45490, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (b) *Perinia* cf. *tumida* (FLMNH 45652, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (c) *Percnon abbreviatum* (FLMNH 45628, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (d) *Dynomene hispida* (FLMNH 69204, Manawai, Hawai&#x02BB;i). Photo credits: Florida Museum of Natural History, licensed [CC BY-NC 3.0](https://creativecommons.org/licenses/by-nc/3.0/)](data/images/4crabs.png){#fig:fig_crabs}

@@

ARMS units were retrieved by encapsulating them in-situ in a 100 μm mesh-lined crate to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate (see [@lerayDNABarcodingMetabarcoding2015] for processing details). All adult brachyuran crabs ($\geq$ 2 mm) were photographed, identified, preserved in 95% ethanol, and retained in a specimen archive at the Hawai&#x02BB;i Institute of Marine Biology. Specimens were initially identified to family or genus using Poore & Ahyong [-@pooreMarineDecapodCrustacea2023] and additionally to species where possible according to expert advice from Scott Godwin (former curator of invertebrate zoology, Bishop Museum) and Gustav Paulay (curator of invertebrate zoology, Florida Museum of Natural History). Where species-level identifications were not possible, taxa were collapsed to family or genus (depending on individual identification). We recognize that the species-level taxonomy of Indo-Pacific brachyurans remains poorly resolved, particularly for abundant groups like the Chlorodiellinae [@lasleyjrIntegrativeSystematicRevision2014; @lasleyjrPhylogeneticRelationshipsUbiquitous2015], and some "species" are known to comprise complexes [e.g. *Perinia* cf. *tumida*, @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009]. However, the prevalence of species complexes and cryptic diversity suggests that our assignments are likely to under- rather than overrepresent the true number of taxa. Thus, any significant patterns we detected likely persist in spite of the potential presence of cryptic diversity. Since our primary concern was categorical diversity of assemblages rather than accurate naming, we have retained species names in cases where complexes are suspected.

@@

```{r tbl-supp-species, results="asis", dependson="taxonomy-validation"}
# make table of all sampled taxa
species_tbl <- cc$crabs %>%
  psmelt() %>%
  # sum counts by island group, depth zone, and species
  group_by(OTU,island_group,hawaiian_name,depth_zone,depth,family,genus,species,complex) %>%
  summarise(n=sum(Abundance)) %>%
  ungroup() %>%
  # keep species we actually found
  filter(n > 0) %>%
  # figure out if a species is exclusive to a depth zone
  group_by(species) %>%
  mutate(zones = str_c(depth_zone,collapse="|")) %>%
  ungroup() %>%
  mutate(depth_exclusive = depth_zone == zones) %>%
  arrange(island_group,hawaiian_name,depth_zone,depth,family,genus,species) %>%
  # join in validated taxonomy
  left_join(crab_taxonomy_validated,by=c("OTU" = "otu")) %>%
  select(hawaiian_name,island_group,depth_zone,depth,family.y,display_species,depth_exclusive,complex=complex.y) %>%
  mutate(display_species = as.character(case_when(
    as.logical(complex) ~ cf_species(display_species),
    .default = display_species
  ))) %>%
  select(-complex) %>%
  rename(island = hawaiian_name,valid_species=display_species) %>%
  mutate(
    across(c(island_group,depth_zone),str_to_sentence),
    depth_exclusive = if_else(depth_exclusive,"Yes","No")
  ) %>%
  rename_with(\(n) {
    n %>%
      str_replace_all("depth$","depth (m)") %>%
      str_replace_all("_"," ") %>%
      str_replace_all("\\.y$","") %>%
      str_replace_all("^display.","") %>%
      str_to_sentence()
  })
print_table(species_tbl,"tbl_supp_species","List of crab taxa detected by island and ARMS unit deployment depth",hdr_case="none",accuracy=1)
```
